### 2018-06-14 paper/referee_reply1.txt
Completed revisions after implementing changes suggested by co-authors

We thank the referee for a thoughtful, expert critique of our paper. We respond to each major and minor comment below, omitting a direct reply to the minor comments when the changes were straightforward to implement.

Comments to the Author
This paper presents size-dependent clustering of SDSS galaxies, and interprets it using a simple model for the dependence of effective radius on halo radius. It is well presented and the results are interesting and useful. I have however a few suggestions/questions that may help to further improve the clarity and strengthen the analysis.

MAJOR:

1) The size-halo connection that the authors advocate, presented in Section 3.2.2, sets each galaxy to be 1% of the size of its halo at the time of Mpeak. The authors state below Eq 3 that this defines a _median_ relation between Reff and Mhalo, around which they assume a log-normal scatter of 0.2 dex. However, it seems that a more direct application of Eq 2 would be to apply it on a halo-by-halo basis, since Rpeak is known for every object in the catalogue. This would specify the full Reff distribution, with scatter deriving from the scatter in Rpeak. Why is this not done? No justification is given for using a log-normal scatter of 0.2 dex.

Reply) The concluding sentence to Section 3.2.2 was indeed confusing; it turns out it was also misleading: in fact, in our analysis we assigned R_1/2 in exactly the fashion the referee described, about each halo's own R_Mpeak. We have changed the language to more accurately reflect the implementation of the model, and thank the referee for pointing out this confusing language. We have also expanded the discussion for our choice of the fiducial level of scatter, as described further in our reply to the next comment.

2) This model is compared to one in which Reff does not correlate with any halo property at fixed stellar mass, but simply depends on M* in a manner that reproduces the M*-Reff relation by construction. Again the authors take 0.2 dex scatter around the median relation, but if one really wanted to reproduce the M*-Reff relation -- including its scatter -- one would take this scatter in bins of M* from the data. Even better would be to drop the assumption of Gaussian scatter altogether and simply assign a size randomly to each model galaxy by sampling the observed Reff distribution at given M*. It appears the authors' choice is motivated by a desire for compatibility with Somerville et al 2017, but it is unclear why this is especially relevant or important here.

Reply) We agree with the referee that drawing directly from P_SDSS(R_1/2 | M*) would improve the fidelity of our model with respect to the SDSS. However, our goal here is actually twofold. On the one hand, we want to identify a model that has sufficient flexibility to capture size-dependent clustering with good accuracy, using SDSS as our validation dataset. On the other hand, we are attempting to build a class of models that can be applied equally well to other datasets, including those may not have the statistics of the SDSS, as is the case in present-day high-redshift surveys. We have added what we hope is a clarifying paragraph on this point towards the end of Section 7.

3) Why use Rvir at Mpeak and not at any other time (say z=0)? I assume this is simply because Mpeak gives the best match to the observations, but this is not explained and hence comes across as somewhat arbitrary. If the result is sensitive to this (as suggested in the final paragraph of Section 5.2), then clustering provides a means to distinguish different choices for halo size in terms of how well they predict Reff. Do the authors know which choices would or would not work? It would be helpful to compare the clustering predicted by one or more alternatives.

Reply) Our ``tidal stripping" model is an attempt to explore the consequences of using Mvir(z=0) to model galaxy size. There are certainly many other possible choices one can make for the halo property used to connect to size, such as the mass at the time of first infall, at the time of last infall, etc. We agree with the referee that a more comprehensive study of the hyper-parameter space is warranted by our results. In particular, we have found one good-fitting model based on R_Mpeak, and several other poor-fitting models based on alternative choices for the galaxy-halo connection. This directly implies that observations of galaxy clustering can indeed discriminate between such choices. Thus we consider a comprehensive study (analogous to the Reddick+13 SHAM study) well-motivated by our work, but beyond the present scope.

4) There should be an uncertainty on w_p in the mock data, i.e. on the lines in Figs. 4-6. Since the model is probabilistic (both the abundance matching step and size assignment have random scatters), different Monte Carlo realisations of the model should give different results. Further, for a given realisation jackknife resampling would give an uncertainty estimate in a similar manner to the SDSS data. I would like to see these uncertainties as bands on the plots, or at least a general estimate of their magnitude. Looking at the bottom-right panel of Fig. 6, for the green line to be consistent with 0 the uncertainty on it must be fairly large, at least at low r; without this uncertainty it is hard to gauge the significance of the difference between the purple line and the data.

Reply) We have added jackknife error bars in the form of shaded bands surrounding wp in the figures comparing models to data.

5) When some of the stellar mass is removed due to stripping, does this not cause the final SMF that's actually used in the model to disagree with the observed input SMF? In the discussion just below Eq 1, it seems that the distribution of M* is the one that matches the observations, not M'*. Or is the SMF that's input to AM modified somehow so that only after this stripping does it match Moustakas et al 2013?

Reply) The implementation of the stellar mass stripping model is indeed crude and to some extent the self-inconsistency pointed out by the referee is inescapable: we are after all not providing a model for the true evolution of post-infall stellar mass, but instead we are grafting an approximate treatment that is intended to approximate the impact of tidal stripping in a statistical sense. That being said, the modification to the SMF is rather small, at or below 10% for the mass ranges relevant to our clustering measurements, and so comparing these two models directly should not be too misleading. We have added a sentence at the end of 3.2.1 calling attention to this detail.

6) A complementary method for studying the size-halo connection is to constrain correlations of galaxy size with halo properties *at fixed stellar mass*, which has been done using both weak lensing (e.g. Sonnenfeld et al 2015, Sonnenfeld & Leauthaud 2017) and galaxy dynamics (e.g. Desmond 2016, 2017). These are means of testing disk models such as Dutton et al 2007 & 2011 which the authors mention. It would be helpful to provide more information on how the Rvir-based model presented here relates to this alternative method. For example, since on average M* is positively correlated with Rpeak by AM, and empirically Reff is positively correlated with M*, then the model in which Reff depends only on M* (Sec 3.2.1) will also positively correlate Reff with Rpeak. It is not clear how this correlation differs from Reff=0.01*Rpeak, or in other words how the latter assumption correlates Reff with halo properties at fixed M*. It would be interesting to see 1) The Reff-Rpeak relation implied by the model in Sec 3.2.1, and/or 2) The correlation of Reff with halo mass or concentration at fixed M* implied by Reff=0.01*Rpeak. Understanding these different approaches will be of great help in devising a unified model for the size-halo connection.

Reply)  We have verified the correlation between R_1/2 and Mvir at fixed M* in present day centrals, and added a paragraph to Section 6 with the suggested references.

7) Final paragraph of Section 6: It has been known for almost 20 years that the scatter in halo spin at fixed mass is larger than the scatter in galaxy size (de Jong & Lacey 2000, Gnedin et al 2007). The forward modelling implications of Reff^{disk} \propto \lambda have already been explored by Desmond & Wechsler 2015, who find it to be problematic for the Tully-Fisher relation as well as the scatter in size. Further, most modern hydrodynamical simulations (including Illustris [Rodriguez-Gomez et al 2017], EAGLE [Sales et al 2009, Desmond et al 2017] and Magneticum [Teklu et al 2015]) correlate Reff little if at all with halo spin. (It would be useful in general to provide a qualitative comparison with hydro sims, which have the potential to test models for the size-halo connection from first principles, and are beginning to do so.) Again it is unclear to me why Somerville et al 2017 should be of any particular importance for the scatter: The scatter in Reff at fixed M* comes straight from the data, and the scatter in lambda at fixed Mhalo (only relevant for angular momentum-based size models, which are barely mentioned here) comes straight from the N-body sim.

Reply) We have added a new sentence that clarifies the purpose of the final paragraph of Section 6, which is to call attention to the methodological similarity of our model to that in Somerville+17. We are interested in applying our forward-modeling techniques to see what insight can be learned about the R_1/2^{disk} \propto \lambda from galaxy clustering, which would be complementary to the knowledge gained by the Desmond & Wechsler 2015 study.

Studying hydro sims with two-point statistics is the subject of ongoing work. We are including, for the referee's interest, a plot showing the results of a recent calculation of the clustering ratios exhibited by Illustris galaxies. We consider these and related results more appropriate for a separate follow-up paper.

We note that although the correlation of size or baryon specific angular momentum with current specific halo angular momentum is found to be weak in current simulations, some simulations show a nearly linear size-virial radius correlation consistent with observations (J. Schaye, private comm.). Also, the observed distribution of galaxy sizes is comparable to that expected in simple spin-based models from the halo spin distribution, when galaxies of *all* types are considered. Thus, the smaller scatter of sizes exhibited by disk galaxies may indicate either that sizes are affected by other factors and not only by spin, or, alternatively, that disk galaxies do not sample the entire spin distribution of halos. Both possibilities are very interesting and motivate future studies, which we believe are outside the scope of this paper.

MINOR:

1) Please provide more details in the main body of the text on the method used to treat orphan subhalos, and in particular that 50% of these are populated with galaxies in the fiducial model. As shown in the Appendix, this has significant impact on the clustering results, so it is important to be up front about the assumptions required. It should also be mentioned in the main text that the AM was done using all subhalos rather than only the ones that survived, as this may lead to substantial differences with alternative AM models that do not model orphans.

2) I recommend replacing M_{halo} everywhere by the specific definition of halo mass which is being referred to, typically M_{peak}. The reader may naturally interpret M_{halo} to mean something else, for example present-day M_{vir}.

Reply) We have changed the text discussing Figure 10 at the end of Section 7. We replaced "M_{halo}" with "M_{peak}", which we agree improves the clarity of the discussion. We are happy to make other analogous changes if there are specific places where the referee finds ambiguity resulting from the use of the term "M_{halo}".

3) DD, DR and RR need to be defined in the calculation of w_p.

4) It may be helpful to normalise the pdfs in Figs 8-9 so that the total area under each curve is 1 (this may already be the case for Fig 8). It is unclear at the moment what scales are used on the y-axes.

5) I suggest that the number of galaxies and stellar mass and redshift range of the datasets used be mentioned in the first paragraph of Section 2, rather than (or in addition to) in Section 4.2. Please justify the z<0.02 cut. Is this done in the simulation as well so that the mock data has the same number of galaxies and redshift distribution as the real data? Also, why does the 10^{11} thresholded sample have the same zmax as the 10^{10.75} sample?

6) Abstract: surely one doesn't "forward model the observations"; this suggests that the observations are an input to the model, which they are precisely not in forward modelling. Perhaps one forward models *to* the observations?

7) I'll let the Scottish grammar pass, but be any misunderstanding on yer ain heids.